\Large\textbf{Preparation GUIDE for Initial-Submission Manuscript}\\~\\ \large\textbf{ Preparation of Papers in Two-Column Format\\ for the ASP-DAC 2014 (\LaTeX2e version)}

### 2013-06-24 background.tex
First version of the paper

\begin{table}[tb]
\centering
\caption{SystemJ kernel statements and their meaning}
\begin{minipage}{8cm}
  \begin{small}
   \begin{tabular}{|c|p{80pt}|}
     \hline                                                                                     
     \textbf{Kernel Statements} & \textbf{Meaning}\\                                            
     \hline                                                                                     
     \hline                                                                                     
     [\textbf{\texttt{input}}] [\textbf{\texttt{output}}]
     [\textbf{\texttt{type}}] \textbf{\texttt{signal}} S & declare signal S\\                                      
     \hline                                                                                     
     \textbf{\texttt{emit}} S [(value)] & broadcast signal S\\                                                    
     \hline                                                                                     
     \textbf{\texttt{present}} (S) \{p\} else \{q\}& do p if S is present, else do q\\                            
     \hline                                                                                     
     \textbf{\texttt{abort}} (S) \{p\} & preempt program p if S is present\\                                      
     \hline                                                                                     
     \textbf{\texttt{suspend}} (S)\{p\} & suspend p if S is present\\                                             
     \hline                                                                                     
     \textbf{\texttt{trap}} (T)\{p\ldots \textbf{\texttt{exit}} T\ldots\} & preempt p if exit is executed\\                         
     \hline                                                                                     
     p\textbf{\texttt{$||$}}q & run p and q in lock-step\\                                                        
     \hline                                                                                     
     p$><$q & run p and q asynchronously\\                                                      
     \hline                                                                                     
     \textbf{\texttt{send}} C([value]) & send a value through C, blocking
     send\\                                                 
     \hline                                                                                     
     \textbf{\texttt{receive}} C() & receive a value through C, blocking
     receive\\
     \hline                                                                                     
     \textbf{\texttt{pause}} & finish a tick and communicate
     with environment\\
     \hline                                                                                     
   \end{tabular}
  \end{small}
   % \footnotetext[1]{\scriptsize Uppercase}
 \end{minipage}
 \label{tab:1}
\end{table}

Table~\ref{tab:1} presents the SystemJ kernel statements used to program
the control flow. The data-flow is programmed in Java. A number of
derived statements also exist (e.g., \texttt{sustain}, \texttt{await},
etc) that make programming reactive systems easier. The \textit{Model of
  Computation} (MoC) of a simple SystemJ is shown in Figure~\ref{fig:2}.

\begin{figure*}[bth]
\centering
\begin{SubFloat}{\label{fig:2a}Simple SystemJ program}%...verbatim subfigure
\begin{minipage}[b]{0.3\linewidth}% a minipage to control the width...
\begin{verbatim}
while(true){
 abort(A){
  while(true)
   pause;
  emit O;
  // do java computation
}
\end{verbatim}%
\end{minipage}%
\end{SubFloat}
\hspace{1.5cm}%
\begin{SubFloat}[Black box]{\label{fig:2b}Ticks in SystemJ}%
\includegraphics[scale=0.5]{moc}
\end{SubFloat}%
\caption{Simple SystemJ example and corresponding MoC}
\label{fig:2}
\end{figure*}

SystemJ consists of entities called clock-domains (CD)s, each running at
their individual pace. Each CD adheres to the \textit{perfect synchrony}
hypothesis, i.e., all statements execute instantaneously in zero
time. Only the \texttt{pause} statement consumes time, just like in
Esterel~\cite{gber931}. Consider the SystemJ program in
Figure~\ref{fig:2a}, it is waiting for an input signal producing logical
ticks (see Figure~\ref{fig:2b}). Once \texttt{A} is received at tick 4,
output \texttt{O} is emitted to the environment instantaneously.

@@

The perfect synchrony hypothesis is ideal for programming the
synchronous sub-set of SystemJ. But, the real-world is not that
forgiving, execution of every \textit{micro-step} in the logical zero
time, requires $\delta$ physical time. Time for a complete reaction to
one or more input signal can thus be summarized as $\Delta = \Sigma
(\delta)$. We call $\Delta$ the reaction time. Obviously depending upon
the amount of computation required, $\Delta$ can vary, hence, in order
to satisfy the implicit restriction posed by the synchrony hypothesis --
no input event can be missed, one needs to calculate the \textit{Worst
  Case Reaction Time} (WCRT) and the resultant WCRT needs to be smaller
than the speed of the fastest input event, else the synchrony hypothesis
is violated. A substantial amount of research for calculating the WCRT
of synchronous programs~\cite{proop10,boldt07,wilhelm08} exists. We
refer the reader to any of these research articles, since WCRT analysis
is not the focus of this paper. The opposite of the WCRT is the
\textit{Best Case Reaction Time} (BCRT), we denote the WCRT and BCRT in
Figure~\ref{fig:2b} using \texttt{W} and \texttt{B}, respectively.

We now present two lemmas with proof sketch (due to lack of space) that
we require for the rest of the paper.

\begin{lemma}
  WCRT and BCRT analysis are invariant to values of conditional
  expressions.
\end{lemma}

\begin{proof}
  The physical time ($\delta$) taken by a conditional expression is at
  best affected by the type rather than the value of the
  conditional. Example, comparing the value of 2 floats might take
  longer than comparing the value of 2 integer types on some given
  platform. But, the time taken by the comparison instruction is
  invariant to the value itself.
\end{proof}

\begin{lemma}
  WCRT and BCRT analysis are invariant to channel communication.
\end{lemma}
\begin{proof}
  LATER ....
\end{proof}

>>>>

\begin{table}[tb]
\centering
\caption{SystemJ kernel statements and their meaning}
\begin{minipage}{8cm}
  \begin{scriptsize}
   \begin{tabular}{|c|p{80pt}|}
     \hline                                                                                     
     \textbf{Kernel Statements} & \textbf{Meaning}\\                                            
     \hline                                                                                     
     \hline                                                                                     
     [\textbf{\texttt{input}}] [\textbf{\texttt{output}}]
     [\textbf{\texttt{type}}] \textbf{\texttt{signal}} S & declare signal S\\                                      
     \hline                                                                                     
     \textbf{\texttt{emit}} S [(value)] & broadcast signal S\\                                                    
     \hline                                                                                     
     \textbf{\texttt{present}} (S) \{p\} else \{q\}& do p if S is present, else do q\\                            
     \hline                                                                                     
     \textbf{\texttt{abort}} (S) \{p\} & preempt program p if S is present\\                                      
     \hline                                                                                     
     \textbf{\texttt{suspend}} (S)\{p\} & suspend p if S is present\\                                             
     \hline                                                                                     
     \textbf{\texttt{trap}} (T)\{p\ldots \textbf{\texttt{exit}} T\ldots\} & preempt p if exit is executed\\                         
     \hline                                                                                     
     p\textbf{\texttt{$||$}}q & run p and q in lock-step\\                                                        
     \hline                                                                                     
     p$><$q & run p and q asynchronously\\                                                      
     \hline                                                                                     
     \textbf{\texttt{send}} C([value]) & send a value through C, blocking
     send\\                                                 
     \hline                                                                                     
     \textbf{\texttt{receive}} C() & receive a value through C, blocking
     receive\\
     \hline                                                                                     
     \textbf{\texttt{pause}} & finish a tick and communicate
     with environment\\
     \hline                                                                                     
   \end{tabular}
  \end{scriptsize}
   % \footnotetext[1]{\scriptsize Uppercase}
 \end{minipage}
 \label{tab:1}
\end{table}

Table~\ref{tab:1} presents the SystemJ kernel statements used to program
the control flow. The data flow is programmed in Java. A number of
derived statements also exist (e.g., \texttt{sustain}, \texttt{await},
etc) that make programming reactive systems easier. The \textit{Model of
  Computation} (MoC) of a simple SystemJ program is shown in
Figure~\ref{fig:2}.

\begin{figure*}[bth]
\centering
\begin{SubFloat}{\label{fig:2a}Simple SystemJ program}%...verbatim subfigure
\begin{minipage}[b]{0.3\linewidth}% a minipage to control the width...
\begin{verbatim}
while(true){
 abort(A){
  while(true)
   pause;
  emit O;
  // do java computation
}
\end{verbatim}%
\end{minipage}%
\end{SubFloat}
\hspace{1.5cm}%
\begin{SubFloat}[Black box]{\label{fig:2b}Logical time in SystemJ}%
\includegraphics[scale=0.4]{moc}
\end{SubFloat}%
\begin{SubFloat}{\label{fig:2c}Mapping to physical time}%
% \includegraphics[scale=0.4]{phy.pdf_t}
\scalebox{0.4}{\input{phy.pdf_t}}
\end{SubFloat}%
\caption{Simple SystemJ example and corresponding MoC}
\label{fig:2}
\end{figure*}

SystemJ consists of entities called clock-domains (CD)s, each running at
their individual pace, interacting with the environment via signals and
with each other via channels and CSP~\cite{pwel05} style
rendezvous. Each CD adheres to the \textit{perfect synchrony}
hypothesis, i.e., all statements execute instantaneously in zero
time. Only the \texttt{pause} statement consumes time, just like in
Esterel~\cite{gber931}. Consider the SystemJ program in
Figure~\ref{fig:2a}, it is waiting for an input signal producing logical
ticks (see Figure~\ref{fig:2b}). Once \texttt{A} is received at tick 4,
output \texttt{O} is emitted to the environment instantaneously.

@@

The perfect synchrony hypothesis is ideal for programming the
synchronous sub-set of SystemJ. But, the real-world is not that
forgiving, execution of every \textit{micro-step} in the logical zero
time, requires $\delta$ physical time. Time for a complete reaction to
one or more input signal can thus be summarized as $\Delta = \Sigma
(\delta)$. We call $\Delta$ the reaction time. Obviously, depending upon
the amount of computation required, $\Delta$ can vary, hence, in order
to satisfy the implicit restriction imposed by the synchrony hypothesis,
no input event can be missed, one needs to calculate the \textit{Worst
  Case Reaction Time} (WCRT) and the resultant WCRT needs to be smaller
than the speed of the fastest input event, else the synchrony hypothesis
is violated. A substantial amount of research for calculating the WCRT
of synchronous programs~\cite{proop10,boldt07,wilhelm08} exists. We
refer the reader to any of these research articles, since WCRT analysis
is not the focus of this paper. The opposite of the WCRT is the
\textit{Best Case Reaction Time} (BCRT), we denote the WCRT and BCRT in
Figure~\ref{fig:2c} using \texttt{W} and \texttt{B},
respectively. Figure~\ref{fig:2c} shows a continuously running physical
clock (analogous to a clock in hardware), the numerical annotations on
the rising edge of the clock mark the logical ticks from
Figure~\ref{fig:2b}. The subscripts \texttt{s} and \texttt{e} for the
numbers show the start and end of the logical ticks, respectively. The
end of a logical tick and the start of the next logical tick happen
together. Figure~\ref{fig:2c} clearly shows the mapping of the logical
time to the physical time. As we can see due to Java computations
(Figure~\ref{fig:2a}) the physical time between logical ticks 4 and 5 is
2 clock cycles, while for the rest of the ticks it is just 1 clock
cycle. This elasticity is inherent (and elegant) to both Esterel and
SystemJ.

% We now present a lemma with proof sketch (due to lack of space) that we
% require for the rest of the paper and is special to SystemJ, being a
% GALS language.

% \begin{lemma}
%   WCRT and BCRT analysis are invariant to values of conditional
%   expressions.
% \end{lemma}

% \begin{proof}
%   The physical time ($\delta$) taken by a conditional expression is at
%   best affected by the type rather than the value of the
%   conditional. Example, comparing the value of 2 floats might take
%   longer than comparing the value of 2 integer types on some given
%   platform. But, the time taken by the comparison instruction is
%   invariant to the value itself.
% \end{proof}

% \begin{lemma}
%   WCRT and BCRT analysis are invariant to channel communication.
% \end{lemma}
% \begin{proof}
%   LATER ....
% \end{proof}

### 2013-06-24 delays.tex
first commit
First version of the paper

\section{Conclusion and future work}
\label{sec:concl-future-work}

>>>>

\section{Introducing real time in SystemJ}
\label{sec:intr-real-time}

We introduce a single \textit{derived} construct called
\mbox{\texttt{delay [M..N]}} in the SystemJ language for real time
control. The resolution of the delay statement is of secondary concern
and is dependent upon the execution platform, without loss of generality
through out this paper we consider the resolution to be always in milli
seconds.

\subsection{Semantics of the delay statement}
\label{sec:semant-delay-stat}

Given a SystemJ program: \texttt{delay [M..N]; p}, where $M \in
\mathbb{Q}^{>0}$ and $N \in \mathbb{Q}^{>0}$, statement $p$ is executed
after real-time $\tau$, such that, $M \leq \tau \leq N$.

We also introduce two variants of the derived statement \texttt{delay}.
\begin{enumerate*}
\item Given a SystemJ program \texttt{delay M; p}, where $M \in
  \mathbb{Q}^{>0}$, statement $p$ is executed after real-time $\tau$,
  such that, $M \leq \tau \leq \infty$. It is important to note that the
  lower bound of the delay construct $M$ is \textit{not} an exact delay,
  but rather the control is allowed to proceed to the next statement
  anytime after the delay of time $M$.
\item Give a SystemJ program \texttt{delay [M..M]; p}, where $M \in
  \mathbb{Q}^{>0}$, statement $p$ is executed after real-time $\tau$,
  such that, $M \leq \tau \leq M$. In this variant, the delay is exact.
\end{enumerate*}

The aforementioned variants are specializations of the general case
\texttt{delay [M..N]}.

\subsection{Rewriting the \texttt{delay} statement}
\label{sec:rewr-delay-stat}

The introduced \texttt{delay} construct is not a kernel construct, but a
\textit{derived} construct built from the kernel constructs
(Table~\ref{tab:1}) in SystemJ. Figure~\ref{fig:3} gives the rewrite of
the \texttt{delay} construct to kernel statements.

\begin{figure}[tb]
    \begin{minipage}{\textwidth}
      \begin{scriptsize}
\begin{verbatim}
trap(T){
 int x = 0;
 while(true){
  x = x + 1;
  if(x == d) exit (T); //wait for "d" ticks
  pause;
 }
}
\end{verbatim}
      \end{scriptsize}
    \end{minipage}
    \caption{The rewrite of \texttt{delay} construct}
    \label{fig:3}
\end{figure}

The fundamental observation is that real-time is converted into logical
time via the \texttt{pause} construct. The rewrite basically maps the
physical notion of time back to the elegant logical notion of time. The
rewrite \textit{delays} a certain number of logical ticks, before
proceeding to the next statement. The number of logical ticks
\texttt{``d''} to \textit{delay} (Figure~\ref{fig:3}) is determined by
the compiler statically at compile time. The value of \texttt{d} is
intricately tied with the WCRT and BCRT of the program and hence the
execution platform.

\subsection{Finding the logical delay \texttt{d}}
\label{sec:find-logic-delay}

\begin{algorithm}[t!]
  \begin{minipage}{1.0\linewidth}
    \SetAlgoLined
    \KwData{WCRT, BCRT, $M \in \mathbb{Q}^{>0}$, $N \in \mathbb{Q}^{>0}$}
    \KwResult{d}
    let $l_1 \leftarrow \lceil \frac{M}{WCRT} \rceil$\;
    let $l_2 \leftarrow \lceil \frac{M}{BCRT} \rceil$\;
    let $u_1 \leftarrow \lfloor \frac{N}{WCRT} \rfloor$\;
    let $u_2 \leftarrow \lfloor \frac{N}{BCRT} \rfloor$\;
    let $F:(l_1,u_1) \rightarrow S_1$\;
    let $F:(l_2,u_2) \rightarrow S_2$\;
    let $D \leftarrow S_1 \cap S_2$\;
    \Return (some $d \in D$)\;
    \caption{Finding the value of \texttt{d}}
    \label{alg:1}
  \end{minipage}
\end{algorithm}

The computation of \texttt{d} is shown in Algorithm~\ref{alg:1}. This
algorithm is carried out for each CD (in SystemJ) or a synchronous
program individually. \textit{The fundamental observation is -- the
  reaction time for each logical tick is elastic -- varying only between
  the BCRT and the WCRT, thus any logical number of ticks \texttt{d}
  that map to the required real-time \texttt{delay} should be chosen in
  such a way that they are invariant to this elasticity.}

Our algorithm takes as input: WCRT, BCRT, and the lower and upper bounds
$M$ and $N$, respectively of the \texttt{delay} construct. We first
divide $M$ and $N$ with the BCRT and WCRT, respectively. This division
gives us the number of individual logical ticks required to delay the CD
(or synchronous program) by the real-time specification. We always
$ceil$ when dividing $M$ and $floor$ when dividing $N$ to make sure that
the resultant values are integers (in domain $\mathbb{N}^{>0}$) and
these functions guarantee that the resultant logical ticks result in
real-time delays between the required range $M-N$. Next, a linear
function $F$ maps these calculated values to a set of equidistant
integer points (values) separated by a unit value -- these points
represent all the logical ticks running at the WCRT and the BCRT,
respectively that satisfy the real-time delay requirements. The
intersection of these two sets gives all the logical ticks that satisfy
the real-time requirements invariant of the reaction time and its
elasticity.

Let us revisit our motivating example to elucidate the algorithm. From
Figure~\ref{fig:1} we know that $M$ is 50ms and $N$ is 100ms,
respectively. Let the WCRT and BCRT be: 13.332ms and 10ms,
respectively. Thus, the algorithm proceeds as follows:

\begin{enumerate*}
\item $l_1 \leftarrow \lceil 50/13.332 \rceil$ and $u_1 \leftarrow
  \lfloor 100/13.332 \rfloor$. $l_1 = 4$ and $u_1 = 7$. We first
  calculate the logical ticks that are always running at the WCRT
  satisfy the required real-time delay.
\item $l_2 \leftarrow \lceil 50/10 \rceil$ and $u_2 \leftarrow
  \lfloor 100/10 \rfloor$. $l_1 = 5$ and $u_1 = 10$. We do the same for
  the BCRT case.
\item $S_1 = \{4,5,6,7\}$ and $S_2 =\{5,6,7,8,9,10\}$. We then map the
  resultant bounds to linear points. Sets $S_1$ and $S_2$ represent
  logical ticks that running at the WCRT and BCRT, respectively always
  satisfy the required real-time constraints.
\item $D = S_1 \cap S_2$, $D = \{5,6,7\}$. Finally, the intersection of
  the two sets gives the set $D$ from which we can pick up any value for
  \texttt{d}.
\end{enumerate*}

The resultant value for \texttt{d} gives the number of logical ticks,
which can run at any physical clock-speed, bounded by the BCRT and the
WCRT of the program and still result in the required real-time delay. We
think this is an elegant solution, because the technique provides
hard-real time guarantee while preserving the essence of synchronous and
GALS programming prescribed by SystemJ and Esterel style
languages. Moreover, this technique considers non maximal-parallelism,
i.e., the delays in logical ticks is calculated after scheduling has
been performed. To our knowledge we are the first to do so.

\subsection{Extending the technique to variants of \texttt{delay}}
\label{sec:extend-tehcn-vari}

\paragraph{The \texttt{delay(M)} construct}
\label{sec:extend-techn-vari}

The \texttt{delay(M)} variant is easily accommodated in the
technique. All one needs to do is find the set $S_1$ and choose a value
from this set.

\paragraph{The \texttt{delay(M..M)} construct}
\label{sec:extend-techn-vari}

The \texttt{delay(M..M)} variant is a little more interesting. Like
before we find sets $S_1$ and $S_2$, and find the intersection of the
two sets to get the value of \texttt{d}. It is possible (and often
likely, as suggested by our experiments in
Section~\ref{sec:experimental-results}) that the resultant set $D$ is
empty (also possible in the case of \texttt{delay(M..N)}, but never
possible in case of \texttt{delay(M)}). In such a case, we automatically
\textit{relax} the upper bound of the \texttt{delay} statement.

\paragraph{Relaxation of the upper real-time bound}
\label{sec:over-appr-relax}

The relaxation algorithm is shown in Algorithm~\ref{alg:2}.

\begin{algorithm}[t!]
  \begin{minipage}{1.0\linewidth}
    \SetAlgoLined
    \KwData{$S_2$, $D$, WCRT}
    \KwResult{d}
    \If {$D = \emptyset$} {
      let $j_{0}$ be the first element of set $S_2$, s.t., $|S_2|=Q$\;
      \ShowLn let $N \leftarrow WCRT \times j_0$\;
      let $d \leftarrow j_0$\;
    }
    \Return d\;
    \caption{Calculating the minimum relaxation of the upper real-time
      bound}
    \label{alg:2}
  \end{minipage}
\end{algorithm}

This algorithm results in the smallest relaxation required for the
real-time delay to be satisfied. The algorithm takes as input the WCRT
and set $S_2$, recall that set $S_2$ represents the logical ticks
required to satisfy the real-time requirement at the BCRT. We take the
very first value from set $S_2$ and multiply it with the WCRT to get the
relaxation $N$. The first element of set $S_2$ is returned as the
logical tick delay \texttt{d}.

The fundamental observation is that we have to delay for a minimum of
$M$ units of real-time, hence, under-approximation is out of
question. We can still over-approximate, but to reduce the resultant
error, we should over-approximate by least possible value, which is the
lower bound of set $S_2$. Thus, the lower bound is considered to be the
only element shared between the two sets $S_1$ and $S_2$ and
accordingly, the upper real-time bound is relaxed by the multiplication
of WCRT and the first element of set $S_2$.

\subsection{Programming using the delay construct}
\label{sec:progr-using-delay}

In this section we provide a number of examples to show the different
types of real-time programming paradigms that can be incorporated into
the GALS (and its sub-set synchronous) programming model.

\subsubsection{Non-deterministic time}
\label{sec:non-determ-time}

A number of programs require non-deterministic timing constructs. One of
them is our motivating example -- the human response time
system. Another is a printer-spooler example borrowed from timed
CSP~\cite{Schneider:1999:CRT:555233}. The spooler and printer need to
synchronize using channels. The printer might be unable to print
depending upon the paper tray, similarly, spooler might take sometime to
send the job depending upon its size. Such real-time constraints can be
modeled in SystemJ as below:

\begin{scriptsize}

\begin{verbatim}
{await (job); delay([2..10]ms);  //SPOOLER CD
 send mid(job) delay (1ms);}
><
{receive mid; delay([1..30]ms); //PRINTER CD
 emit print(#mid);}
\end{verbatim}
\end{scriptsize}

\subsubsection{Timeout}
\label{sec:timeout}

There are many instances when one would like to wait on an input from
environment for only a specified amount of time. This can be programmed
as:
    \begin{scriptsize}
\begin{verbatim}
// timeout after 1ms.
trap(T){
 {await(A);}||{delay(1ms);exit(T); }
}
\end{verbatim}
    \end{scriptsize}

\subsubsection{Periodic reactions}
\label{sec:periodic-reactions}

A reaction, or a whole CD, can be programmed to run periodically like
so:

\begin{scriptsize}

\begin{verbatim}
{
delay (1ms); emit S; //do something
}
\end{verbatim}
\end{scriptsize}
A periodic reaction (or CD) requires special consideration. Since the
delay statement is converted into \texttt{pause} constructs. One should
not introduce extra \texttt{pause} constructs when building periodic
reactions (or CDs). This is essential since, introducing pauses would
introduce more logical ticks.

\subsubsection{Interaction of preemption and delays}
\label{sec:inter-preempt-delays}

Preemption plays an important role in reactive languages. One needs to
carefully consider the interplay of \texttt{delay} semantics with the
preemption semantics of reactive languages. Other attempts at
incorporating delays (using external timers) have only been partially
successful, because of the complex interplay between real-time and
preemption. Consider the simple example below, which models real-time
using external timers as in~\cite{rsh94}.

\begin{scriptsize}

\begin{verbatim}
suspend{S} {
 emit START_TIMER(10); await (TIMER); 
 emit O1;
}
\end{verbatim}
\end{scriptsize}

As identified in~\cite{Bourke2009a} the \texttt{suspend} does not play
well with the external timer. The above program sends a signal to an
external timer and waits for 10ms to pass by. Consider what happens when
signals \texttt{S} and \texttt{TIMER} occur together, the \texttt{await}
statement is never executed and hence, we enter a deadlock. Such
problems are completely avoided in our technique, because we convert the
real-time delays into logical delays (\texttt{pause} constructs), which
bode well with preemption.

\subsubsection{Interaction of channel communication and delays}
\label{sec:inter-chann-comm}

Channels are an addition in the SystemJ language. Like interaction of
preemption and delays, conversion of rela-time delays to logical ticks
also bodes well with channel rendezvous, because the semantics of
interaction are well defined~\cite{amal10}. More importantly, we need to
consider the interplay of channel communication and WCRT/BCRT
analysis. Since, channel communication does not stop logical time
(see~\cite{amal10}) WCRT/BCRT are unaffected by channel
communication. The response time to input signals \textit{is} though!
But, we are unconcerned with the response time, analysis in this paper
and it remains a future research avenue.

### 2013-06-24 introduction.tex
First version of the paper

Reactive languages~\cite{gber05,amal10} are a class of programming
languages used for designing and implementing \textit{reactive} systems,
which continuously respond to input from their environment. These
languages have been successfully used in programming a plethora of
systems such as fly by wire in Airbus~\cite{eairbus}, security
surveillance systems~\cite{amal121}, etc. Usually the working
environment of these reactive systems is also real-time. Yet, these
languages do not support describing real-time as first class language
constructs. For example, one cannot describe a simple real-time delay
\texttt{delay 0.2 ms} in these languages. These languages are based on
formal semantics, essential for the verification of functional
correctness properties of the developed programs, but leave
non-functional properties such as timing behavior as an implementation
detail~\cite{ringler00,boldt07}. Arguably rightly so, because physical
time cannot be incorporated without information about the underlying
execution platform. But, one can still reason in terms of
\textit{logical} time. These languages support discrete \textit{logical}
clock rather than a continuous clock. The period of the discrete logical
clock (also referred to as a \textit{logical tick}) is determined by the
responsiveness of the program to the input signals and as opposed to a
discrete \textit{physical} clock is elastic. Although this timing model
has worked well for these languages in designing discrete systems there
is obviously a need to introduce continuous time since often
implementation models are in the continuous time
domain~\cite{DBLP:journals/pieee/SifakisTY03}.

@@

\begin{enumerate}
\item The real-time is modeled in the $\mathbb{Q}^{>0}$ domain of
  numbers, i.e., non-negative rational numbers.
\item We consider the property of \textit{non maximal-parallelism},
  i.e., we consider scheduling on constrained resources to be an
  integral part of the solution. To our knowledge we are the very first
  to do so.
\item Our solution does not change the functional or timing semantics of
  GALS and its subset synchronous programs, in fact, our solution does
  not require one to change the compiler at all.
\item Our solution gives complete control of under and over
  approximation of real-time to the system designer.
\item We are the first to allow specification of both deterministic and
  non-deterministic real-time.
\end{enumerate}

>>>>

Reactive languages~\cite{gber05,amal10} are a class of programming
languages used for designing and implementing \textit{reactive} systems,
which continuously respond to input from their environment. These
languages have been successfully used in programming a plethora of
systems such as fly by wire in Airbus~\cite{eairbus}, security
surveillance systems~\cite{amal121}, etc. Usually the working
environment of these reactive systems is also real-time. Yet, these
languages do not support describing real-time as first class language
constructs. For example, one cannot describe a simple real-time delay
\texttt{delay 0.2 ms} in these languages. These languages are based on
formal semantics, essential for the verification of functional
correctness properties of the developed programs, but leave
non-functional properties such as timing behavior as an implementation
detail~\cite{ringler00,boldt07}. Arguably, rightly so, because physical
time cannot be incorporated without information about the underlying
execution platform. But, one can still reason in terms of
\textit{logical} time. These languages support discrete \textit{logical}
clock rather than a continuous clock. The period of the discrete logical
clock (also referred to as a \textit{logical tick}) is determined by the
responsiveness of the program to the input signals and as opposed to a
discrete \textit{physical} clock, is elastic. Although this timing model
has worked well for these languages in designing discrete systems, there
is obviously a need to introduce continuous time since often
implementation models are in the continuous time
domain~\cite{DBLP:journals/pieee/SifakisTY03}.

@@

\begin{enumerate*}
\item Real-time is modeled in the $\mathbb{Q}^{>0}$ domain of numbers,
  i.e., non-negative rational numbers.
\item We consider the property of \textit{non maximal-parallelism},
  i.e., we consider scheduling on constrained resources to be an
  integral part of the solution. To our knowledge we are the very first
  to do so.
\item Our solution does not change the functional or timing semantics of
  GALS and its subset synchronous programs, in fact, our solution does
  not require one to change the mid-end, the back-end or the related
  optimization phases of the compiler at all.
\item We are the first, to our knowledge, to allow specification of both
  deterministic and non-deterministic real-time in GALS and synchronous
  languages.
\end{enumerate*}

### 2013-06-24 motivation.tex
First version of the paper

% system{
%  interface{
%   input signal TOUCH;
%   output signal GREEN_LIGHT;
%   output char channel S;
%   input char channel R;
%  }
\begin{figure}[t!]
% \begin{center}
\begin{minipage}{5cm}
\begin{verbatim}
{ // Clock-domain 1
while(1) {
  trap(TRAP){
   // Abort if user touches the screen
   abort(TOUCH){
    {sustain GREEN_LIGHT;}
    || // synchronous parallel operator
    {
    //exit after any 
    // where from 0.5 to 1 ms.
    /*delay [0.5..1]ms;*/
    exit(TRAP); 
    }
   } do {send S(1);} // Send 1 if touched 
  } do {send S(0);} // Else Send 0
  pause; // only construct to consume time
 }
}
>< // asynchronous operator
{ // Clock-domain 2
  while(1){ receive S;}
  // do something with S
}
\end{verbatim}
\end{minipage}
\caption{An example human reaction time collection system (HRTCS)}
% }
% \end{center}
\label{fig:1}
\end{figure}

Consider a simple machine used to test human reaction time and collect
this data for future research purposes. The machine switches on a green
light on a touch screen for anywhere between \texttt{0.5ms} to
\texttt{1ms}, if the user can touch this green light then she is
successful and an integer \texttt{1} is sent to a data-base server, if
unsuccessful a \texttt{0} is sent instead, this way a database of human
responsiveness can easily be generated.

@@

The SystemJ program is divided into two clock-domains, the first
clock-domain continuously displays a green light (\texttt{GREEN\_LIGHT}
signal) on the touch screen sensor and waits for the user to respond. If
the user is able to touch the screen within the specified time of
\texttt{0.5ms} to \texttt{1ms} a positive response is sent to the second
clock-domain, which is interacting with data-base (via CSP style channel
communication via \texttt{S}) server in Java. The synchronous parallel
operator \texttt{||} along with the \texttt{trap-exit} (exception
mechanism) is used to make sure that the green light is switched on for
only a specified delay. The 2 clock-domains are implemented using the
asynchronous parallel \texttt{><} operator. All these constructs are
described in~\cite{amal10}.

This SystemJ example is \textit{incomplete}, the \texttt{delay}
specification, which models the non-deterministic delay between
\texttt{0.5ms} and \texttt{1ms} does not work with the SystemJ compiler!
In fact, there is no way to include such real-time specifications in the
language itself.

>>>>

% system{
%  interface{
%   input signal TOUCH;
%   output signal GREEN_LIGHT;
%   output char channel S;
%   input char channel R;
%  }
\begin{figure}[t!]
% \begin{center}
\begin{minipage}{5cm}
  \begin{scriptsize}

\begin{verbatim}
{ // Clock-domain 1
while(1) {
  trap(TRAP){
   // Abort if user touches the screen
   abort(TOUCH){
    {sustain GREEN_LIGHT;}
    || // synchronous parallel operator
    {
    //exit after any 
    // where from 50 to 100 ms.
    /*delay [50..100]ms;*/
    exit(TRAP); 
    }
   } do {send S(1);} // Send 1 if touched 
  } do {send S(0);} // Else Send 0
  pause; // only construct to consume time
 }
}
>< // asynchronous operator
{ // Clock-domain 2
  while(1){ receive S;}
  // do something with S
}
\end{verbatim}
  \end{scriptsize}
\end{minipage}
\caption{An example human response time collection system (HRTCS)}
% }
% \end{center}
\label{fig:1}
\end{figure}

Consider a simple machine used to test human response time and collects
this data for future research purposes. The machine switches on a green
light on a touch screen for anywhere between \texttt{50ms} to
\texttt{100ms}, if the user can touch this green light then she is
successful and an integer \texttt{1} is sent to a database server, if
unsuccessful a \texttt{0} is sent instead, this way a database of human
responsiveness can easily be generated.

@@

The SystemJ program is divided into two clock-domains, the first
clock-domain continuously displays a green light (\texttt{GREEN\_LIGHT}
signal) on the touch screen sensor and waits for the user to respond. If
the user is able to touch the screen within the specified time of
\texttt{50ms} to \texttt{100ms} a positive response is sent to the
second clock-domain, which is interacting with a database (via CSP style
channel communication via \texttt{S}) server in Java. The synchronous
parallel operator \texttt{||} along with the \texttt{trap-exit}
(exception mechanism) is used to make sure that the green light is
switched on for only a specified delay. The 2 clock-domains are
implemented using the asynchronous parallel \texttt{><} operator. All
these constructs are described in~\cite{amal10} and
Section~\ref{sec:background}.

This SystemJ example is \textit{incomplete}, the \texttt{delay}
specification, which models the non-deterministic delay between
\texttt{0.5ms} and \texttt{100ms} does not work with the SystemJ compiler!
In fact, there is no way to include such real-time specifications in the
language itself.

### 2013-06-24 related_work.tex
first commit
First version of the paper

\section{Introduction}
\label{sec:introduction}

>>>>

\section{Related work}
\label{sec:related-work}

A small number of studies have been performed when it comes to
integrating real-time directly as a programming construct in GALS (and
its subset synchronous) reactive languages. The most prominent works in
this area are by Shyamsundar~\cite{rsh94} and Bourke et
al.~\cite{Bourke2009a}. Shyamsundar incorporates real-time using
external timers in \textit{Communicating Reactive Processes} (CRP),
which like SystemJ is an extension of synchronous language Esterel to
asynchorny. But, as mentioned in Section~\ref{sec:inter-preempt-delays},
external timers do not interact well with preemption constructs in these
languages. Bourke et al. do a much better job of introducing real-time
as first class constructs in the Esterel language. They like us provide
real-time \texttt{delay} as first class Esterel programming constructs,
and translate them into Esterel kernel constructs. But, unlike us they
do not translate delays into \texttt{pause} constructs
directly. Instead, logical ticks are generated by using abstract notion
of \texttt{event} and \texttt{sample} platform dependent timers. This
notion makes it a complex and inflexible solution since the number of
platform timers with certain resolutions need to be determined and
present on the system for the solution to be realizable. Moreover,
non-deterministic real-time delays and integration with non-maximal
parallelism is not studied at all by them. Other works such as
Quartz~\cite{glog02} also incorporate delay statements in Esterel using
\texttt{pause} constructs, but again, non-maximal parallelism,
non-deterministic delays are not studied. Moreover, Quartz is targeted
at studying timing properties using model checkers rather than
implementation, like us.
